---
title: "DataSummary"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## LPBO Data Summary 

This code was written to facilitate the download and cleaning of LPBO annual data for the purpose of creating annual outputs for reporting.

### Load Libraries 

```{r}

#If using NatureCounts for the first time, install the package
#install.packages("remotes")
#remotes::install_github("BirdsCanada/naturecounts")

library(naturecounts)
library(tidyverse)

```

### Create Directories 

```{r}

#Create data folder and output folders in working directory
if(!dir.exists("Data")) dir.create("Data")
if(!dir.exists("Output")) dir.create("Output")

#Assign directories
out.dir <- "Output/"
data.dir <- "Data/"

```

### Download Data

You will need a NatureCounts account and have requested permission to access the data through the R package.

```{r}

year=2024 #specify the year you want to download
req_id = XXXXXX #specify the quest ID that was sent to you via email once the data download was approved
ID = "YOURUSERNAME"

LPBO<-nc_data_dl(collections ="CMMN-DET-LPBO", years= year, fields_set = "extended", username = ID, info = "Date for end of your summary", request_id = req=id)


#save a copy of the data in your `Data` directory
write.csv(LPBO, paste(data.dir, "LPBO_", year, ".csv"), row.names = FALSE)
LPBO<-read.csv(paste(data.dir, "LPBO_", year, ".csv"))

```


### Data Preparation

```{r}

#Select only the columns you need
dat <- LPBO %>% select(species_id, CommonName, YearCollected, MonthCollected, DayCollected, JulianDay, record_id, SiteCode, ObservationCount, ObservationDescriptor, ObservationCount2, ObservationDescriptor2, ObservationCount3, ObservationDescriptor3, ObservationCount4, ObservationDescriptor4, ObservationCount5, ObservationDescriptor5)

#Effort measurement need transferred to the BMDE.

#Summarize the total number of species, sum of species band/census/DET, and record_id by station
sum1<- dat %>% group_by(SiteCode) %>% summarise(TotalUniqueSpecies = n_distinct(species_id), TotalBand=sum(as.numeric(ObservationCount4), na.rm = TRUE), TotalCensus=sum(as.numeric(ObservationCount3), na.rm=TRUE), TotalDET=sum(as.numeric(ObservationCount), na.rm=TRUE), TotalRecords = n_distinct(record_id))

#Include a final row in the above table what sums the columns for all stations
sum2<-sum1 %>% summarise(TotalUniqueSpecies = max(TotalUniqueSpecies), TotalBand=sum(TotalBand), TotalCensus=sum(TotalCensus), TotalDET=sum(TotalDET), TotalRecords = sum(TotalRecords)) %>% mutate(SiteCode="All") %>% select(SiteCode, TotalUniqueSpecies, TotalBand, TotalCensus, TotalDET, TotalRecords)

#Combine the two tables
sum<-rbind(sum1, sum2)

#Create a spring and fall variable based on the month, where spring runs for May to July and fall runs from August to November
dat$Season<-ifelse(dat$MonthCollected %in% c(5,6,7), "Spring", ifelse(dat$MonthCollected %in% c(8,9,10,11), "Fall", "Other")) 
#remove the "Other" season
dat<-dat %>% filter(Season %in% c("Spring", "Fall"))

#Create a summary table for spring and fall seperate, that summarize the total number of species, sum of species band/census/DET, and record_id by station, and the number of days each station was active

sum3<- dat %>% group_by(SiteCode, Season) %>% summarise(TotalUniqueSpecies = n_distinct(species_id), TotalBand=sum(as.numeric(ObservationCount4), na.rm = TRUE), TotalCensus=sum(as.numeric(ObservationCount3), na.rm=TRUE), TotalDET=sum(as.numeric(ObservationCount), na.rm=TRUE), TotalRecords = n_distinct(record_id), DaysActive=n_distinct(JulianDay))

#Include a final row in the above table what sums the columns for all stations
sum4<-sum3 %>% group_by(Season) %>% summarise(TotalUniqueSpecies = max(TotalUniqueSpecies), TotalBand=sum(TotalBand), TotalCensus=sum(TotalCensus), TotalDET=sum(TotalDET), TotalRecords = sum(TotalRecords), DaysActive=sum(DaysActive)) %>% mutate(SiteCode="All") %>% select(SiteCode, Season, TotalUniqueSpecies, TotalBand, TotalCensus, TotalDET, TotalRecords, DaysActive)

#Combine the two tables
sum_spfall<-rbind(sum3, sum4)


```

Total for the year
-	Total volunteer/person hours
-	Total species and forms reported
-	Total number of individuals banded
-	Total number of species and forms banded
For Spring and Fall, separately….
-       Total volunteer/person hours
-       Total species and forms reported
-       Total number of individuals banded
-       Total number of species and forms banded
-       Total station-days of migration monitoring (for Old Cut, Breakwater, and Tip separately)
-       Total net hours (mist net)
-       Total trap hours (J-trap and Ground trap separately)
-       Overall catch rate (birds/100 hours)
-       Total number of visitors to Old Cut
